---
title: "PS1_Yannan_Niu"
format: html
editor: visual
---

# Problem set 1

## Problem 1

### a. Import data and rename columns

```{r}
#| code-fold: true
# Import data
wine <- read.csv("/Users/nynn/Library/CloudStorage/OneDrive-Umich/Umich course/2024_Fall/Stats 506/Stats_506_PS1/wine/wine.data", header = FALSE)

# Change names according to wine.name
names(wine)
names <- c("Class", "Alcohol", "Malic acid", "Ash",
            "Alcalinity of ash",  "Magnesium", "Total phenols", "Flavanoids",
            "Nonflavanoid phenols", "Proanthocyanins", "Color intensity", 
           "Hue", "OD280/OD315 of diluted wines", "Proline")
names(wine) <- names
head(wine)

```

###  b. Check data

```{r}
#| code-fold: true
# Check wine classes
table(wine$Class)
```

### c.1 What is the correlation between alcohol content and color intensity?

```{r}
#| code-fold: true
plot(x=wine$Alcohol, y=wine$`Color intensity`)
round(cor(x=wine$Alcohol, y=wine$`Color intensity`), 2)

```

**The correlation between alcohol content and color intensity is 0.55.**

### c.2 Which class has the highest correlation? Which has the lowest?

```{r}
#| code-fold: true
class(wine$Class)

## Creat a function to calculate the correlation within each class
cor_class <- function(class) {
  cor <- cor(x=wine$Alcohol[wine$Class == class], 
             y=wine$`Color intensity`[wine$Class == class])
  print(paste("Class", as.character(class), ": ", as.character(round(cor, 2))))
  return(cor)
}

class <- unique(wine$Class)
sapply(class, cor_class)
```

**Class 1 has the highest correlation of 0.41 and Class 2 has the lowest correlation of 0.27.**

### c.3 What is the alcohol content of the wine with the highest color intensity?

```{r}
#| code-fold: true
wine$Alcohol[which.max(wine$`Color intensity`)]
```

**The alcohol content of the wine with the highest color intensity is 14.34%.**

### c.4 What percentage of wines had a higher content of proanthocyanins compare to ash?

```{r}
#| code-fold: true
n_true <- nrow(wine[(wine$Proanthocyanins > wine$Ash), ] )
n_total <- nrow(wine)
prop_true <- 100* round(n_true/n_total,4)
print(paste("Percentage: ", prop_true, "%"))
```

**8.43% wines had a higher content of proanthocyanins compare to ash.**

### d. Create a table for average values

```{r}
#| code-fold: true
## Creat a function to calculate the mean for each variable and within classes

calculte_mean <- function() {
  class <- c(0, 1, 2, 3)
  
  colmean_class <- function(class) {
    # Select class
    if (class == 0) {
    selected_class <- wine[ , 2:14]
     } else {
    selected_class <- wine[ , 2:14][wine$Class == class, ]}
    # Calculate
    mean <- colMeans(selected_class)
    return(mean)
  }
  
  m <- sapply(class, colmean_class)
  # Rename cols
  colnames(m) <- c("Overall", "Class 1", "Class 2", "Class 3") 
  
  return(m)
}

## Use the function
calculte_mean()

```

### e. T test for three classes

```{r}
#| code-fold: true
# Subest dataset for different classes
phe_class_1 <- wine$`Nonflavanoid phenols`[wine$Class == 1]
phe_class_2 <- wine$`Nonflavanoid phenols`[wine$Class == 2]
phe_class_3 <- wine$`Nonflavanoid phenols`[wine$Class == 3]

ttest_12 <- t.test(phe_class_1, phe_class_2)
ttest_13 <- t.test(phe_class_1, phe_class_3)
ttest_23 <- t.test(phe_class_2, phe_class_3)
ttest_12$statistic
ttest_12$p.value < 0.05
ttest_13$statistic
ttest_13$p.value < 0.05
ttest_23$statistic
ttest_23$p.value < 0.05

```

**The t-tests result in large t-values and p-values less than 0.05 indicate significant differences in phenol levels across the three wine classes. These results suggest that phenol levels are not uniform across the classes, with each group showing distinct mean differences.**

## Problem 2

### a. Import data

```{r}
#| code-fold: true
AskAManager <- read.csv("/Users/nynn/Library/CloudStorage/OneDrive-Umich/Umich course/2024_Fall/Stats 506/Stats_506_PS1/AskAManager.csv")
```

### b. Rename variables

```{r}
#| code-fold: true
names(AskAManager)
  
new_names <- c("ID", "Time", "Age", "Industry", "Job", "Job_add",
                        "Salary", "Compensation", "Currency", "Currency_add",
                        "Income_add", "Country", "State", "City", "Work_Year_All", 
                        "Work_Year_Field", "Education", "Gender", "Race")
length(new_names)

names(AskAManager) <- new_names
```

### c. Restrict data with USD

```{r}
#| code-fold: true
print(paste("The number of observation before restriction: ", nrow(AskAManager)))
AskAManager_USD <- AskAManager[AskAManager$Currency == "USD", ]
print(paste("The number of observation after restriction: ", nrow(AskAManager_USD)))
```

### d. Restrict age with 18

```{r}
#| code-fold: true
unique(AskAManager$Age)
unique(AskAManager$Work_Year_All)
unique(AskAManager$Work_Year_Field)

# Factor variables
AskAManager$Age_new <- as.numeric(factor(AskAManager$Age, levels = c("under 18", "18-24", "25-34", "35-44", "45-54", "55-64", "65 or over")))
AskAManager$Work_Year_All_new <- as.numeric(factor(AskAManager$Work_Year_All, levels = c("1 year or less", "2 - 4 years", "5-7 years", "8 - 10 years", "11 - 20 years", "21 - 30 years", "31 - 40 years", "41 years or more")))
AskAManager$Work_Year_Field_new <- as.numeric(factor(AskAManager$Work_Year_Field, levels = c("1 year or less", "2 - 4 years", "5-7 years", "8 - 10 years", "11 - 20 years", "21 - 30 years", "31 - 40 years", "41 years or more")))

# Function for the maximum work years
Work_All_most_calculate<- function(Age_new) {
  if (as.numeric(Age_new) == 1) {
    Work_Year_All_most = 0
  } else if(as.numeric(Age_new) == 2) {
    Work_Year_All_most = 3
  } else {
    Work_Year_All_most = Age_new + 2
  }
  return(Work_Year_All_most)
}
AskAManager$Max_Work_Year_All <- sapply(AskAManager$Age_new, Work_All_most_calculate)

# Employee must be (a) larger than 18, 
                #  (b) length should be less than years since 18, 
                #  (c) field work time should be less than all work time 
AskAManager_Age <- AskAManager[AskAManager$Age_new > 1 
                    & AskAManager$Work_Year_All_new <= AskAManager$Max_Work_Year_All
                    & AskAManager$Work_Year_All_new >= AskAManager$Work_Year_Field_new, ]

print(paste("The number of observation before restriction: ", nrow(AskAManager)))
print(paste("The number of observation after restriction: ", nrow(AskAManager_Age)))

```

### e. Restrict salaries

**I used the 95% confidence interval as the threshold for valid salary. I assumed the salary distributes normally because of the large sample size.**

```{r}
summary(AskAManager$Salary)

mean <- mean(AskAManager$Salary)
sd <- sd(AskAManager$Salary)
n <- length(AskAManager$Salary)
se <- sd/sqrt(n)

# Assume normal distribution, z = 1.96 for 95%
lower <- mean - 1.96 * se
upper <- mean + 1.96 * se

AskAManager_Salary <- AskAManager[AskAManager$Salary >= lower & AskAManager$Salary <= upper, ]
print(paste("The number of observation before restriction: ", nrow(AskAManager)))
print(paste("The number of observation after restriction: ", nrow(AskAManager_Salary)))

```

### f. Statistical association between education and salary

```{r}
AskAManager$Education <- as.factor(AskAManager$Education)

model <- lm(Salary ~ Education + Work_Year_All + Work_Year_Field, data = AskAManager)

summary(model)
```

**After controlling years of experience with the total work years and the work year in a specific field, there is no significant association between education and salary.**
